Ontology Matching with word2vec

### 2016-06-17 Paper/Draft.tex
Added full running example to paper

 \subsubsection{Example}
 An example of an ontology or a knowledge base of mice and men would be the following:
 A mouse consists of the following body parts:
 \begin{itemize}
 \item Head
 \item Torso
 \item Whiskers
 \item Tail
 \item Front paws
 \item Hind paws
 \end{itemize}

@@

 How should the wildly differing demands the applications place on the ontology matching systems be satisfied by a single of a few systems? When merging ontologies on one end of the Aristotle-Plato spectrum, often a completely different approach is required than an ontology on the other side. Sometimes the matching needs to be very fast, for example when the ontologies are used in a search query which the user expects to be done in milliseconds.

@@

@@

  The most basic \textbf{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{ear lobe} and \emph{ear} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.

@@

@@

  Another way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in {appropriate section}, and selecting from the different node senses based on which is closest to the context average.

@@

@@

  % maybe ref too\endcomment
  This means we can for example have the context be (previous word, next word) and those words will be treated differently. Similarly we can separate the edge and node labels and treat them differently. This ensures that the model will find any relationship between the edge and the node if it exists and will take this into account. For example, the system might use just the information 'isMarriedTo' to infer that someone is a human, but may infer that 'isMarriedTo' in combination with a man usually refers to a woman.
  %  Ja doe maar een ander voorbeeld ivm gelijkheid e.d.

@@

  It is hard to classify the proposed algorithm in terms of the conventional ontology alignment method classes. It is not purely string matching, since it takes into account information from the context. However, it is also not structural, as it can work without a seed alignment and uses information other than the structure as well. It is not logical 
  %  Goeie woord?
  since it does not use inference. Nor is it terminological, which looks at dictionaries or other ontologies to find matches. However, it does use other ontologies (or text corpora) implicitly, so if it must be grouped it would be terminological.

@@

@@

  Word embedding models need to be trained on a large corpus. These corpora need to cover the concepts that are contained in the ontology, but also need to be large enough to build good embeddings. Since the ontologies themselves do not necessarily contain enough examples to embed the concepts properly, I need to use ontologies that are likely to contain the concepts I want to align.

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense \textbf{Word2vec}.

>>>>

 \subsubsection{Example} \label{example}
 An example of an ontology or a knowledge base of mice and men would be the following:
 A mouse consists of the following body parts:
 \begin{itemize}
 \item Head
 \item Torso
 \item Whiskers
 \item Tail
 \item Front paws
 \item Hind paws
 \end{itemize}

@@

 How should the wildly differing demands the applications place on the ontology matching systems be satisfied by a single or just a few systems? When merging ontologies on one end of the Aristotle-Plato spectrum, often a completely different approach is required than an ontology on the other side. Sometimes the matching needs to be very fast, for example when the ontologies are used in a search query which the user expects to be done in milliseconds.

@@

 The explanations refer back to the example in Section \ref{example}

@@

  The most basic \textbf{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{whisker} and \emph{moustache} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.

@@

  For example, since the \emph{front paws} and \emph{arms} are both attached to the \emph{torso}, their context vectors will be similar as well.

@@

  Another way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in {appropriate section}, and selecting from the different node senses based on which is closest to the context average. %TODO appropriate section

@@

  In the case of the mouse Ontology, the \emph{head} is above the \emph{torso} so \emph{isAbove} will be added to the context of \emph{head}. The same goes for the \emph{human head}, which will make them more similar.

@@

  %TODO appropriate section
  %TODO maybe ref too\endcomment
  This means we can for example have the context be (previous word, next word) and those words will be treated differently. Similarly we can separate the edge and node labels and treat them differently. This ensures that the model will find any relationship between the edge and the node if it exists and will take this into account. For example, \emph{isBelow} and \emph{torso} will combine into a vector that is similar to \emph{tail} whereas \emph{isAbove} and \emph{torso} combine into a vector that is similar to \emph{head} since the edge and node labels are combined in the model, rather than considered separately.

@@

  It is hard to classify the proposed algorithm in terms of the conventional ontology alignment method classes. It is not purely string matching, since it takes into account information from the context. However, it is also not structural, as it can work without a seed alignment and uses information other than the structure as well. It is not semantic (or logical) since it does not use inference. Nor is it terminological, which looks at dictionaries or other ontologies to find matches. However, it can use dictionaries and other ontologies to improve performance.

@@

  An example of a hot start is that \emph{mouse torso} is already matched with \emph{human torso}. Then it is likely that their neighbours are also related. Since the \emph{front paws} and \emph{arms} are next to the \emph{torso}, they are more likely to be matched now. In the case of \textbf{Node2vec} they are more likely to be matched, since their neighbourhoods are the same (the only neighbour of both \emph{arms} and \emph{paws} is \emph{connectedTo torso}).

@@

  Word embedding models need to be trained on a large corpus. These corpora need to cover the concepts that are contained in the ontology, but also need to be large enough to build good embeddings. Since the ontologies themselves do not necessarily contain enough examples to embed the concepts properly, I need to use ontologies that are likely to contain the concepts I want to align. In the case of the mice and men ontologies, anatomical or medical ontologies can be useful, since they may list more anatomical relationships or common relationships (for example \emph{whisker} and \emph{moustache} are related to \emph{hair}).

@@

 \subsection{Implementation}
 The foundation the algorithm has been build and tested on is Factorie, described in \ref{factorie}. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense \textbf{Word2vec}.
